### 2024-09-24 paper/paper_Li.qmd
Add files via upload

---
title: "Shelter Demand in Toronto: A Study of Capacity and Utilization(2024)"
author:
  - YiZhuo Li
thanks: "A GitHub Repository containing all data, R code, and other files used in this investigation is located here: <https://github.com/eason1218/Toronto_Shetler.git>"
date: today
date-format: long
abstract: "This paper analyzes 2024 data on shelter occupancy and capacity in Toronto to assess the city's ability to meet the needs of its homeless population. The findings reveal that most shelters operate at or near full capacity, with some reaching 100% occupancy. This highlights a growing demand for emergency shelter services and raises concerns about the city's capacity to support vulnerable groups. Understanding these trends is critical for developing policies that can better address homelessness and improve the allocation of social services in urban settings."
format: pdf
number-sections: true
# The toc parameter used to create a table of contents was obtained from
# Quarto documentation:
# https://quarto.org/docs/reference/formats/pdf.html
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#install.packages("tidyverse")
#install.packages("opendatatoronto")
#install.packages("here")
library(tidyverse)
library(opendatatoronto)
library(here)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

Homelessness remains a significant social issue in many urban centers, and Toronto is no exception. The city’s shelter system serves as a critical component in providing relief to its homeless population. However, the increasing demand for these services, coupled with limited capacity, has highlighted potential gaps in Toronto’s ability to address this growing crisis. A comprehensive understanding of shelter utilization is essential to inform policies that can better support vulnerable populations.

This paper analyzes Toronto’s 2024 shelter occupancy and capacity data, focusing on the relationship between bed availability and occupancy rates. The study aims to evaluate whether the current shelter infrastructure meets the needs of homeless individuals and families in the city. The investigation identifies occupancy relative to the date, place, and size of the shelter. Despite the critical role shelters play, there remains a clear gap in understanding how different factors affect shelter capacity and utilization. This study seeks to fill that gap by providing a detailed data analysis and identifying potential areas for improvement in Toronto’s shelter system.

The findings show that most shelters operate near or at full capacity, underscoring the urgent need for expanded services. These results are crucial for informing future social policies aimed at mitigating homelessness, particularly in rapidly growing urban environments like Toronto.

With respect to the structure of this paper, @sec-data includes an overview of the data and @sec-model show the results of model; @sec-discussion includes a discussion of the results; and @sec-appendix includes supplementary information.

# Data {#sec-data}

## Overview {#sec-data-overview}

The dataset used in this analysis is from The "Daily Shelter & Overnight Service Usage" page , this website offers daily data of the occupancy and capacity of Toronto’s shelter system, including how many shelter beds are being used and the total available capacity, updated everyday except on weekends and holidays.

The variables or measurements included in this analysis are "Bed Based Capacity", which refer to Occupancy is measured by the number of beds, typically for programs with shared sleeping areas; "Room Based Capacity" refer to ccupancy is measured by the number of rooms, usually for family programs or hotel-based programs where rooms are not shared between households.

Using the R programming language [@citeR], the `tidyverse` [@thereferencecanbewhatever] package were used to simulate the dataset and generate tests for it. The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@thereferencecanbewhatever] packages were then applied in order to download the raw Toronto Public shelter dataset. Next, the `tidyverse` package [@thereferencecanbewhatever] was used to clean the raw dataset and test the cleaned dataset.

## Measurement {#sec-data-Measurement}
It focuses on capturing daily occupancy, vacancy, and capacity data through a system snapshot taken at 4 a.m. each day. This approach ensures consistent, up-to-date information on the availability and usage of shelter services.

\newpage

# Model {#sec-model}

```{r}
#| echo: false
#| warning: false
#| message: false

#### Loading the Cleaned Dataset ####

#outbreak_cleaned_data <- read_csv(here::here("outputs/data/cleaned_data.csv"))
```

After loading the dataset using the R programming language [@citeR] and the `here` package [@citehere], the `tidyverse` [@thereferencecanbewhatever] package was used to generate graphs. In doing so, R code was adapted from [@tellingstorieswithdata].

```{r}
#| label: fig-occupancy-rates
#| fig-cap: Bed and Room Occupancy Rates in Shelters Over Time
#| echo: false
#| warning: false
#| message: false

marriage_licence_packages <- search_packages("Daily Shelter & Overnight Service Occupancy & Capacity")
# Retrieve a list of all resources (datasets) available within the found package
marriage_licence_resources <- marriage_licence_packages %>%
  list_package_resources()
# Display the list of resources available in the "Marriage Licence Statistics" package

marriage_licence_statistics <- marriage_licence_resources[5,] %>%
  get_resource()
analysis_data = marriage_licence_statistics

invisible(Sys.setlocale("LC_TIME", "C"))

# Assuming analysis_data is already loaded and contains the necessary fields
analysis_data$OCCUPANCY_DATE <- as.Date(analysis_data$OCCUPANCY_DATE, format="%Y-%m-%d")

# Group by date and calculate the average bed and room occupancy rates
daily_occupancy_rate <- analysis_data %>%
  group_by(OCCUPANCY_DATE) %>%
  summarise(
    Bed_Occupancy_Rate = mean(OCCUPANCY_RATE_BEDS, na.rm = TRUE),
    Room_Occupancy_Rate = mean(OCCUPANCY_RATE_ROOMS, na.rm = TRUE)
  )

# Plot the occupancy rate trend
ggplot(daily_occupancy_rate, aes(x = OCCUPANCY_DATE)) +
  geom_line(aes(y = Bed_Occupancy_Rate, color = "Bed Occupancy Rate"), size = 1) +
  geom_line(aes(y = Room_Occupancy_Rate, color = "Room Occupancy Rate"), size = 1, linetype = "dashed") +
  labs(title = "", x = "Date", y = "Occupancy Rate (%)") +
  theme_minimal() +
  scale_color_manual(values = c("Bed Occupancy Rate" = "blue", "Room Occupancy Rate" = "red")) +
  theme(legend.title = element_blank())

```
@fig-occupancy-rates shows the occupancy rates for beds and rooms in the shelter system over time. The blue line represents the bed occupancy rate, while the red line represents the room occupancy rate. Overall, the room occupancy rate (red) is higher and more stable, staying close to 100%. In contrast, the bed occupancy rate (blue) shows more fluctuations, generally ranging between 98% and 99%.

\newpage

```{r}
#| label: fig-occupancy-capital-relat
#| fig-cap: Scatter Plot of Bed Capacity vs. Occupancy Rate
#| echo: false
#| warning: false
#| message: false
# Convert OCCUPANCY_RATE_BEDS to numeric (in case of any issues during import)
analysis_data$OCCUPANCY_RATE_BEDS <- as.numeric(analysis_data$OCCUPANCY_RATE_BEDS)

# Calculate the average occupancy rate per location
avg_occupancy_rate <- aggregate(OCCUPANCY_RATE_BEDS ~ LOCATION_NAME, analysis_data, mean, na.rm = TRUE)

# Sort the data by occupancy rate
avg_occupancy_rate <- avg_occupancy_rate[order(avg_occupancy_rate$OCCUPANCY_RATE_BEDS), ]

# Plot the bar chart with smaller text
ggplot(avg_occupancy_rate, aes(x = reorder(LOCATION_NAME, OCCUPANCY_RATE_BEDS), y = OCCUPANCY_RATE_BEDS)) +
  geom_bar(stat = "identity", fill = "skyblue", width = 0.6) +  # Use a more pleasing color
  coord_flip() +  # Flip to make it horizontal
  labs(title = "", x = "Location", y = "Occupancy Rate (%)") +
  theme_minimal() +  # Clean theme for better aesthetics
  theme(
    plot.title = element_text(size = 10, face = "bold", hjust = 0.5),  # Smaller title size
    axis.title = element_text(size = 9),  # Smaller axis title size
    axis.text.x = element_text(size = 7),  # Smaller size of x-axis text
    axis.text.y = element_text(size = 3.8),  # Smaller size of y-axis text (locations)
    panel.grid.major.y = element_blank()  # Remove y-axis grid lines for better readability
  ) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))  # Ensure bars stay within 0-100% limits
```

```{r}
#| label: fig-occupancy-capital-123
#| fig-cap: Scatter Plot of Bed Capacity vs. Occupancy Rate
#| echo: false
#| warning: false
#| message: false
# Clean the data by removing NA values in CAPACITY_ACTUAL_BED and OCCUPANCY_RATE_BEDS
scatter_data <- na.omit(analysis_data[, c("CAPACITY_ACTUAL_BED", "OCCUPANCY_RATE_BEDS")])

# Create a scatter plot using ggplot2
ggplot(scatter_data, aes(x = CAPACITY_ACTUAL_BED, y = OCCUPANCY_RATE_BEDS)) +
  geom_point(alpha = 0.7, size = 1) +  # Set point size to 1 and transparency to 0.7
  labs(title = "", 
       x = "Actual Bed Capacity", 
       y = "Occupancy Rate (Beds)") +
  theme_minimal()  # Use minimal theme
```

@fig-occupancy-capital-relat displays the average occupancy rate across various shelters. The X-axis represents the occupancy rate percentage (Occupancy Rate %), ranging from 0% to 100%, while the Y-axis lists the shelter locations (Location), showing the names of several shelters. Each horizontal blue bar corresponds to the average occupancy rate of a specific shelter. Most shelters have an occupancy rate close to 100%, indicating that their beds are almost fully or completely occupied. A few shelters have occupancy rates below 75%, showing that they have relatively lower bed utilization.

\newpage

@fig-occupancy-capital-123 shows the relationship between bed capacity and occupancy rate across shelters. The X-axis represents bed capacity, and the Y-axis represents occupancy rate (%). Most shelters have an occupancy rate close to 100%, with bed capacities mainly below 100, indicating that many shelters are near full capacity. In some cases, shelters even exceed their available beds.

\newpage

# Discussion{#sec-discussion}

## First discussion point {#sec-first-point}

The data in @fig-occupancy-rates highlights the consistently high demand for shelter services in Toronto. The stable room occupancy rate near 100% suggests that rooms, typically reserved for families or those requiring more privacy, are in greater demand. This is likely due to the more permanent or semi-permanent sheltering needs of these groups. The high demand for such spaces underscores the need for expanded resources that cater to families and vulnerable individuals who prioritize long-term stability.

On the other hand, the fluctuations in bed occupancy rates imply a more dynamic use of shelter services. Beds, often allocated in shared living spaces, are likely occupied by individuals facing short-term or emergency shelter needs. Factors such as weather conditions (e.g., colder months driving up demand), economic changes, or local crises could be contributing to this variability. This fluctuation in bed occupancy indicates that the shelter system must remain flexible to accommodate rapid changes in demand.

From a policy perspective, the findings call for a dual approach to addressing homelessness. On one hand, the stability in room occupancy suggests the need for more permanent housing solutions, especially for families. On the other hand, the fluctuating bed occupancy points to the importance of scalable, emergency response systems that can expand in times of crisis, ensuring that shelters are adequately prepared for demand surges.

## Second discussion point

@fig-occupancy-capital-relat paints a sobering picture of Toronto's shelter system, with most facilities operating at or near full capacity. This level of demand points to a systemic issue that extends beyond shelter availability to broader socioeconomic challenges. The high occupancy rates reflect not just a lack of shelters but underline key drivers of homelessness, such as inadequate affordable housing, mental health challenges, unemployment, and other economic pressures.

The shelters operating below 75% capacity may indicate inefficiencies in resource allocation or barriers to accessibility. These disparities raise critical questions: Are some shelters harder to access due to location or transportation issues? Are the services provided in certain shelters of lower quality, causing individuals to avoid them? These issues suggest that future interventions should not only focus on increasing shelter capacity but also on improving the quality and accessibility of services across all shelters to ensure a more equitable distribution of resources.

## Third discussion point

@fig-occupancy-capital-123 reveals that larger shelters—those with capacities of over 100 beds—tend to operate at higher efficiency, maintaining near 100% occupancy. This observation suggests the presence of economies of scale in shelter management, where larger shelters are able to provide more comprehensive services and cater to a broader range of needs. These shelters likely benefit from centralized resources, better staffing, and more organized operations, which improve their ability to serve more individuals effectively.

However, this observation also raises concerns about the experience of those residing in large shelters. While efficiency is important, overcrowded conditions can negatively impact the quality of life for shelter residents. Future policy should strike a balance between scaling up shelters to meet demand and ensuring that the conditions in these shelters remain humane and conducive to recovery and long-term stability. There should also be investment in smaller shelters that cater to specific groups, such as women, youth, or individuals with mental health challenges, providing them with targeted support that large, generalized shelters might not be able to offer.

## Weaknesses and next steps

Weaknesses  
There may be significant geographic disparities in the availability and occupancy rates of shelters across different areas of Toronto, which could lead to imbalanced resource allocation[@jadidzadeh2023homelessshelter]. Some areas with higher homeless populations may have fewer shelters, placing additional strain on these facilities and causing them to operate at full capacity or beyond. Conversely, areas with a larger number of shelters may see lower utilization rates, meaning resources may not be fully utilized in those regions. Future research should delve deeper into the geographical distribution of shelters in Toronto and assess how geographic factors influence the ability of homeless individuals to access shelter services.

Next steps  
Shelter utilization is influenced not only by the availability of beds and rooms but also by the conditions of the shelters, the quality of services provided, and the specific needs and preferences of homeless individuals. By conducting interviews with shelter users and staff, it would be possible to uncover the complex reasons behind fluctuating occupancy rates[@ranasinghe2006zoning]. Future research should incorporate qualitative data, such as user feedback and service quality assessments, to develop more comprehensive solutions to homelessness.

\newpage

\appendix

# Appendix {#sec-appendix}

## Dataset and Graph Sketches {#sec-appendix-sketches}

Sketches depicting both the desired dataset and the graphs generated in this analysis are available in the GitHub Repository.

## Data Cleaning {#sec-appendix-cleaning}

The data cleaning process involved filtering out some of the columns from the raw dataset and renaming some of the data entries for clarity and simplicity.

## Attribution Statement {#sec-appendix-attribution}

"Contains information licensed under the Open Government Licence – Toronto" [@Shelterdataset].

\newpage

\newpage

# References {#sec-references}

### 2024-09-24 scripts/03-test_data.R
Update 03-test_data.R

@@ -1,15 +1,23 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests clean_data data no obvious statistic mistakes. 
+# Author: YiZhuo Li
+# Date: Today
+# Contact: liyizhuo.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Already have clean data
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Test data ####
+data <- read_csv("C:\\Users\\EasonLi\\Downloads\\Toronto-main\\Toronto-main\\data\\analysis_data_clean.csv")
+
+#test 1 date
+date_format_test <- all(grepl("^\\d{4}-\\d{2}-\\d{2}$", data$OCCUPANCY_DATE))
+
+#test 2 rate > 0 
+rate_beds_test <- all(data$OCCUPANCY_RATE_BEDS[!is.na(data$OCCUPANCY_RATE_BEDS)] > 0)
+rate_rooms_test <- all(data$OCCUPANCY_RATE_ROOMS[!is.na(data$OCCUPANCY_RATE_ROOMS)] > 0)
+
+#test 3 capicity > 0
+capacity_test <- all(data$`Actual Bed Capacity`[!is.na(data$`Actual Bed Capacity`)] > 0)

### 2024-09-24 scripts/01-download_data.R
Update 01-download_data.R

@@ -11,18 +11,17 @@
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Download data ####
 marriage_licence_packages <- search_packages("Daily Shelter & Overnight Service Occupancy & Capacity")
-# Retrieve a list of all resources (datasets) available within the found package
+
 marriage_licence_resources <- marriage_licence_packages %>%
   list_package_resources()
-# Display the list of resources available in the "Marriage Licence Statistics" package
 
 marriage_licence_statistics <- marriage_licence_resources[5,] %>%
   get_resource()
 analysis_data = marriage_licence_statistics
-# 使用 write.csv() 函数保存数据
+
 write.csv(analysis_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data.csv", row.names = FALSE)
 

### 2024-09-24 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -28,5 +28,5 @@ shelter_cleaned_data <-
 
   
 #### Saving the Cleaned Dataset ####
-# 使用 write.csv() 
+
 write.csv(shelter_cleaned_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data_clean.csv", row.names = FALSE)

### 2024-09-24 paper/paper_Li.qmd
Update paper_Li.qmd

@@
 ---
-title: "Shelter Demand in Toronto: A Study of Capacity and Utilization(2024)"
+title: "Capacity and Utilization analysis of Toronto shelter(2024)"
 author:
@@
 The dataset used in this analysis is from The "Daily Shelter & Overnight Service Usage" page , this website offers daily data of the occupancy and capacity of Toronto’s shelter system, including how many shelter beds are being used and the total available capacity, updated everyday except on weekends and holidays.
+The variables or measurements included in this analysis are "Bed Based Capacity", which refer to Occupancy is measured by the number of beds, typically for programs with shared sleeping areas; "Room Based Capacity" refer to ccupancy is measured by the number of rooms, usually for family programs or hotel-based programs where rooms are not shared between households.
 
-The variables or measurements included in this analysis are "Bed Based Capacity", which refer to Occupancy is measured by the number of beds, typically for programs with shared sleeping areas; "Room Based Capacity" refer to ccupancy is measured by the number of rooms, usually for family programs or hotel-based programs where rooms are not shared between households.
+Another dataset called "toronto-shelter-system-flow" also includes data about Toronto shelters, but it is more focused on the characteristics of people living in shelters, not the macro data so I did not choose it.
 
@@
 
-# Model {#sec-model}
+# Result {#sec-model}
 

### 2024-09-26 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,19 +1,27 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates table of shelter in Toronto
+# Author: YiZhuo Li
+# Date: Today
+# Contact: liyizhuo.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: No
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(dplyr)
 
-#### Simulate data ####
-# [...ADD CODE HERE...]
-
+set.seed(1)
 
 
+simulated_data <- data.frame(
+  OCCUPANCY_DATE = seq(as.Date('2024-01-01'), by = "day", length.out = 100), 
+  OCCUPANCY_RATE_BEDS = sample(seq(80, 100, by = 0.5), 100, replace = TRUE),  
+  OCCUPANCY_RATE_ROOMS = sample(seq(80, 100, by = 0.5), 100, replace = TRUE),  
+  LOCATION_NAME = sample(c('COSTI Hotel Program Dixon', 'COSTI Reception Centre', 'COSTI Uptown Hotel Program'), 
+                         100, replace = TRUE),  
+  Actual_Bed_Capacity = sample(seq(5, 50, by = 1), 100, replace = TRUE) 
+)
+
+
+head(simulated_data)

### 2024-09-26 paper/paper_Li.qmd
Update paper_Li.qmd

@@
 
-#install.packages("tidyverse")
-#install.packages("opendatatoronto")
-#install.packages("here")
+# install.packages("tidyverse")
+# install.packages("opendatatoronto")
+# install.packages("here")
 library(tidyverse)
@@
 # Introduction
-
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
 
@@
 
-Using the R programming language [@citeR], the `tidyverse` [@thereferencecanbewhatever] package were used to simulate the dataset and generate tests for it.
+The project is structure base on Rohan's principes [@rohan], and used the **R** programming language [@citeR], the `tidyverse` [@thereferencecanbewhatever] package were used to simulate the dataset and generate tests for it.
 The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@thereferencecanbewhatever] packages were then applied in order to download the raw Toronto Public shelter dataset.
@@
 
-#outbreak_cleaned_data <- read_csv(here::here("outputs/data/cleaned_data.csv"))
+# outbreak_cleaned_data <- read_csv(here::here("outputs/data/cleaned_data.csv"))
 ```
@@
 
-marriage_licence_statistics <- marriage_licence_resources[5,] %>%
+marriage_licence_statistics <- marriage_licence_resources[5, ] %>%
   get_resource()
-analysis_data = marriage_licence_statistics
+analysis_data <- marriage_licence_statistics
 
@@
 # Assuming analysis_data is already loaded and contains the necessary fields
-analysis_data$OCCUPANCY_DATE <- as.Date(analysis_data$OCCUPANCY_DATE, format="%Y-%m-%d")
+analysis_data$OCCUPANCY_DATE <- as.Date(analysis_data$OCCUPANCY_DATE, format = "%Y-%m-%d")
 
@@
   theme(legend.title = element_blank())
-
 ```
@@
   geom_point(alpha = 0.7, size = 1) + # Set point size to 1 and transparency to 0.7
-  labs(title = "", 
+  labs(
+    title = "",
     x = "Actual Bed Capacity",
-       y = "Occupancy Rate (Beds)") +
+    y = "Occupancy Rate (Beds)"
+  ) +
   theme_minimal() # Use minimal theme

### 2024-09-26 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -15,12 +15,14 @@ set.seed(1)
 
 
 simulated_data <- data.frame(
-  OCCUPANCY_DATE = seq(as.Date('2024-01-01'), by = "day", length.out = 100), 
-  OCCUPANCY_RATE_BEDS = sample(seq(80, 100, by = 0.5), 100, replace = TRUE),  
-  OCCUPANCY_RATE_ROOMS = sample(seq(80, 100, by = 0.5), 100, replace = TRUE),  
-  LOCATION_NAME = sample(c('COSTI Hotel Program Dixon', 'COSTI Reception Centre', 'COSTI Uptown Hotel Program'), 
-                         100, replace = TRUE),  
-  Actual_Bed_Capacity = sample(seq(5, 50, by = 1), 100, replace = TRUE) 
+  OCCUPANCY_DATE = seq(as.Date("2024-01-01"), by = "day", length.out = 100),
+  OCCUPANCY_RATE_BEDS = sample(seq(80, 100, by = 0.5), 100, replace = TRUE),
+  OCCUPANCY_RATE_ROOMS = sample(seq(80, 100, by = 0.5), 100, replace = TRUE),
+  LOCATION_NAME = sample(c("COSTI Hotel Program Dixon", "COSTI Reception Centre", "COSTI Uptown Hotel Program"),
+    100,
+    replace = TRUE
+  ),
+  Actual_Bed_Capacity = sample(seq(5, 50, by = 1), 100, replace = TRUE)
 )
 
 

### 2024-09-26 scripts/01-download_data.R
Update 01-download_data.R

@@ -11,17 +11,18 @@
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-
+# [...UPDATE THIS...]
 
 #### Download data ####
 marriage_licence_packages <- search_packages("Daily Shelter & Overnight Service Occupancy & Capacity")
-
+# Retrieve a list of all resources (datasets) available within the found package
 marriage_licence_resources <- marriage_licence_packages %>%
   list_package_resources()
+# Display the list of resources available in the "Marriage Licence Statistics" package
 
-marriage_licence_statistics <- marriage_licence_resources[5,] %>%
+marriage_licence_statistics <- marriage_licence_resources[5, ] %>%
   get_resource()
-analysis_data = marriage_licence_statistics
+analysis_data <- marriage_licence_statistics
 
-write.csv(analysis_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data.csv", row.names = FALSE)
+write.csv(analysis_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/raw_data.csv", row.names = FALSE)
 

### 2024-09-26 scripts/01-download_data.R
Update 01-download_data.R

@@ -11,7 +11,6 @@
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Download data ####
 marriage_licence_packages <- search_packages("Daily Shelter & Overnight Service Occupancy & Capacity")

### 2024-09-26 scripts/03-test_data.R
Update 03-test_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Tests clean_data data no obvious statistic mistakes. 
+# Purpose: Tests clean_data data no obvious statistic mistakes.
 # Author: YiZhuo Li
 # Date: Today
 # Contact: liyizhuo.li@mail.utoronto.ca
@@ -12,12 +12,12 @@ library(tidyverse)
 
 data <- read_csv("C:\\Users\\EasonLi\\Downloads\\Toronto-main\\Toronto-main\\data\\analysis_data_clean.csv")
 
-#test 1 date
+# test 1 date
 date_format_test <- all(grepl("^\\d{4}-\\d{2}-\\d{2}$", data$OCCUPANCY_DATE))
 
-#test 2 rate > 0 
+# test 2 rate > 0
 rate_beds_test <- all(data$OCCUPANCY_RATE_BEDS[!is.na(data$OCCUPANCY_RATE_BEDS)] > 0)
 rate_rooms_test <- all(data$OCCUPANCY_RATE_ROOMS[!is.na(data$OCCUPANCY_RATE_ROOMS)] > 0)
 
-#test 3 capicity > 0
+# test 3 capicity > 0
 capacity_test <- all(data$`Actual Bed Capacity`[!is.na(data$`Actual Bed Capacity`)] > 0)

### 2024-09-26 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -12,21 +12,23 @@
 library(tidyverse)
 
 #### Cleaning the Dataset ####
-shelter_raw_data <- read_csv("C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data.csv")
+shelter_raw_data <- read_csv("C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/raw_data.csv")
 
-shelter_cleaned_data <- 
+shelter_cleaned_data <-
   shelter_raw_data |>
-  
-  select(`OCCUPANCY_DATE`, `OCCUPANCY_RATE_BEDS`, 
-         `OCCUPANCY_RATE_ROOMS`, `LOCATION_NAME`, `CAPACITY_ACTUAL_BED`) |>
-  
-  rename(`OCCUPANCY_DATE` = `OCCUPANCY_DATE`,
-         `OCCUPANCY_RATE_BEDS` = `OCCUPANCY_RATE_BEDS`,
-         `OCCUPANCY_RATE_ROOMS` = `OCCUPANCY_RATE_ROOMS`,
-         `LOCATION_NAME` = `LOCATION_NAME`,
-         `Actual Bed Capacity` = `CAPACITY_ACTUAL_BED`)
+  select(
+    `OCCUPANCY_DATE`, `OCCUPANCY_RATE_BEDS`,
+    `OCCUPANCY_RATE_ROOMS`, `LOCATION_NAME`, `CAPACITY_ACTUAL_BED`
+  ) |>
+  rename(
+    `OCCUPANCY_DATE` = `OCCUPANCY_DATE`,
+    `OCCUPANCY_RATE_BEDS` = `OCCUPANCY_RATE_BEDS`,
+    `OCCUPANCY_RATE_ROOMS` = `OCCUPANCY_RATE_ROOMS`,
+    `LOCATION_NAME` = `LOCATION_NAME`,
+    `Actual Bed Capacity` = `CAPACITY_ACTUAL_BED`
+  )
 
-  
+
 #### Saving the Cleaned Dataset ####
-
+# 使用 write.csv()
 write.csv(shelter_cleaned_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data_clean.csv", row.names = FALSE)

### 2024-09-27 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -18,17 +18,12 @@ shelter_cleaned_data <-
   shelter_raw_data |>
   select(
     `OCCUPANCY_DATE`, `OCCUPANCY_RATE_BEDS`,
-    `OCCUPANCY_RATE_ROOMS`, `LOCATION_NAME`, `CAPACITY_ACTUAL_BED`
+    `OCCUPANCY_RATE_ROOMS`, `LOCATION_NAME`, `CAPACITY_ACTUAL_BED`, `SECTOR`
   ) |>
   rename(
-    `OCCUPANCY_DATE` = `OCCUPANCY_DATE`,
-    `OCCUPANCY_RATE_BEDS` = `OCCUPANCY_RATE_BEDS`,
-    `OCCUPANCY_RATE_ROOMS` = `OCCUPANCY_RATE_ROOMS`,
-    `LOCATION_NAME` = `LOCATION_NAME`,
     `Actual Bed Capacity` = `CAPACITY_ACTUAL_BED`
   )
 
 
 #### Saving the Cleaned Dataset ####
-# 使用 write.csv()
 write.csv(shelter_cleaned_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data_clean.csv", row.names = FALSE)

### 2024-09-27 paper/paper_Li.qmd
Update paper_Li.qmd

@@
 # install.packages("here")
+library(opendatatoronto)
 library(tidyverse)
-library(opendatatoronto)
 library(here)
+library(knitr)
 ```
@@
 
+```{r}
+#| label: tbl-sample
+#| echo: false
+#| warning: false
+#| message: false
+# Load the data from the specified path
+data <- read.csv("C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data_clean.csv")
+
+# Display the first 5 rows with all columns
+kable(head(data, 5),
+  caption = "Table: Sample of cleaned occupancy data",
+  col.names = c(
+    "Occupancy Date", "Occupancy Rate (Beds)", "Occupancy Rate (Rooms)",
+    "Location Name", "Actual Bed Capacity", "Sector"
+  )
+)
+```
+
+@tbl-sample show a sample of cleaned occupancy data, Since the statistical record formats of different shelters are different, each row contains different information marked as NA, but this does not negattive affect the analysis.
 \newpage
@@
 #| message: false
-# Convert OCCUPANCY_RATE_BEDS to numeric (in case of any issues during import)
-analysis_data$OCCUPANCY_RATE_BEDS <- as.numeric(analysis_data$OCCUPANCY_RATE_BEDS)
+#|
+# 转换 'OCCUPANCY_RATE_BEDS' 为数值型，过滤掉缺失值
+analysis_data <- analysis_data %>%
+  mutate(OCCUPANCY_RATE_BEDS = as.numeric(OCCUPANCY_RATE_BEDS)) %>%
+  filter(!is.na(OCCUPANCY_RATE_BEDS))
 
-# Calculate the average occupancy rate per location
-avg_occupancy_rate <- aggregate(OCCUPANCY_RATE_BEDS ~ LOCATION_NAME, analysis_data, mean, na.rm = TRUE)
+# 计算每个位置和 sector 的平均占用率
+avg_occupancy_rate <- analysis_data %>%
+  group_by(SECTOR, LOCATION_NAME) %>%
+  summarize(OCCUPANCY_RATE_BEDS = mean(OCCUPANCY_RATE_BEDS, na.rm = TRUE)) %>%
+  arrange(SECTOR, OCCUPANCY_RATE_BEDS)
 
-# Sort the data by occupancy rate
-avg_occupancy_rate <- avg_occupancy_rate[order(avg_occupancy_rate$OCCUPANCY_RATE_BEDS), ]
+sector_colors <- c("Men" = "#66c2a5", "Mixed Adult" = "#fc8d62", "Women" = "#8da0cb", "Youth" = "#e78ac3")
 
-# Plot the bar chart with smaller text
-ggplot(avg_occupancy_rate, aes(x = reorder(LOCATION_NAME, OCCUPANCY_RATE_BEDS), y = OCCUPANCY_RATE_BEDS)) +
-  geom_bar(stat = "identity", fill = "skyblue", width = 0.6) + # Use a more pleasing color
-  coord_flip() + # Flip to make it horizontal
-  labs(title = "", x = "Location", y = "Occupancy Rate (%)") +
-  theme_minimal() + # Clean theme for better aesthetics
+ggplot(avg_occupancy_rate, aes(x = reorder(LOCATION_NAME, OCCUPANCY_RATE_BEDS), y = OCCUPANCY_RATE_BEDS, fill = SECTOR)) +
+  geom_bar(stat = "identity", width = 0.6) +
+  coord_flip() +
+  labs(title = "", x = "", y = " (%)") +
+  theme_minimal() +
   theme(
-    plot.title = element_text(size = 10, face = "bold", hjust = 0.5), # Smaller title size
-    axis.title = element_text(size = 9), # Smaller axis title size
-    axis.text.x = element_text(size = 7), # Smaller size of x-axis text
-    axis.text.y = element_text(size = 3.8), # Smaller size of y-axis text (locations)
-    panel.grid.major.y = element_blank() # Remove y-axis grid lines for better readability
+    plot.title = element_text(size = 0, face = "bold", hjust = 0.5),
+    axis.title = element_text(size = 0),
+    axis.text.x = element_text(size = 5),
+    axis.text.y = element_text(size = 4),
+    panel.grid.major.y = element_blank(),
+    legend.position = "none"
   ) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) # Ensure bars stay within 0-100% limits
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
+  facet_wrap(~SECTOR, scales = "free_y", ncol = 2) +
+  scale_fill_manual(values = sector_colors)
 ```
-
 ```{r}
@@
 
-@fig-occupancy-capital-relat displays the average occupancy rate across various shelters.
+@fig-occupancy-capital-relat displays the average occupancy rate across various shelters, categorized by sector: Men, Women, Mixed Adult, and Youth.
 The X-axis represents the occupancy rate percentage (Occupancy Rate %), ranging from 0% to 100%, while the Y-axis lists the shelter locations (Location), showing the names of several shelters.
@@
 
-@fig-occupancy-capital-relat paints a sobering picture of Toronto's shelter system, with most facilities operating at or near full capacity.
-This level of demand points to a systemic issue that extends beyond shelter availability to broader socioeconomic challenges.
-The high occupancy rates reflect not just a lack of shelters but underline key drivers of homelessness, such as inadequate affordable housing, mental health challenges, unemployment, and other economic pressures.
+@fig-occupancy-capital-relat reveals that across all sectors—Men, Women, Mixed Adults, and Youth.
+Both show a sobering picture of Toronto's shelter system, with most facilities operating at or near full capacity—this level of demand points to a systemic issue that extends beyond shelter availability to broader socioeconomic challenges.
+The high occupancy rates reflect not just a lack of shelters but also underline key drivers of homelessness, such as inadequate affordable housing, mental health challenges, unemployment, and other economic pressures.
 
@@
 
-The data cleaning process involved filtering out some of the columns from the raw dataset and renaming some of the data entries for clarity and simplicity.
+The data cleaning process involved selecting some columns of useful: occupancy date, bed and room occupancy rates, location name, actual bed capacity, and sector.
+To improve clarity, the column "CAPACITY_ACTUAL_BED" is renamed to "Actual Bed Capacity".
 

### 2024-09-27 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -11,19 +11,18 @@
 library(tidyverse)
 library(dplyr)
 
-set.seed(1)
-
-
+# Simulate data
 simulated_data <- data.frame(
   OCCUPANCY_DATE = seq(as.Date("2024-01-01"), by = "day", length.out = 100),
   OCCUPANCY_RATE_BEDS = sample(seq(80, 100, by = 0.5), 100, replace = TRUE),
   OCCUPANCY_RATE_ROOMS = sample(seq(80, 100, by = 0.5), 100, replace = TRUE),
-  LOCATION_NAME = sample(c("COSTI Hotel Program Dixon", "COSTI Reception Centre", "COSTI Uptown Hotel Program"),
-    100,
-    replace = TRUE
-  ),
-  Actual_Bed_Capacity = sample(seq(5, 50, by = 1), 100, replace = TRUE)
+  LOCATION_NAME = sample(c("COSTI Hotel Program Dixon", "COSTI Reception Centre", "COSTI Uptown Hotel Program"), 100, replace = TRUE),
+  Actual_Bed_Capacity = sample(seq(5, 50, by = 1), 100, replace = TRUE),
+  Sector = sample(c("Men", "Women", "Mixed Adult", "Youth"), 100, replace = TRUE)
 )
 
-
 head(simulated_data)
+
+# Save the simulated data to a CSV file
+write.csv(simulated_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/simulate_data/simulated_shelter_data.csv", row.names = FALSE)
+

### 2024-09-27 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -25,4 +25,3 @@ head(simulated_data)
 
 # Save the simulated data to a CSV file
 write.csv(simulated_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/simulate_data/simulated_shelter_data.csv", row.names = FALSE)
-

### 2024-09-27 other/datasheet/datasheet_template.qmd
Add files via upload

---
title: "Datasheet for 'Toronto_Shelter'"
author: 
  - YiZhuo Li
thanks: "A GitHub Repository containing all data, R code, and other files used in this investigation is located here: <https://github.com/eason1218/Toronto_Shetler.git>"
date: today
date-format: "D MMMM YYYY"
abstract: "The 'Toronto_Shelter' dataset tracks shelter usage in Toronto, including occupancy rates, available beds, and shelter types, updated daily by the city's Housing and Homelessness Services department. It's self-contained, lacks sensitive information, and supports policy-making and resource allocation for homelessness services. Available through an API on the City of Toronto’s official site, the dataset can be used to analyze trends and predict shelter demand but should not be used for individual-level analysis. Missing values marked as “NA” may require consideration during analysis."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The census aims to monitor the capacity, utilization, and turnover in Toronto’s shelter system to understand the changing population of people experiencing homelessness and to inform policy decisions.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The data is collected and managed by the City of Toronto, specifically by the Housing and Homelessness Services department.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The initiative is typically funded through the city’s budget allocations and relevant housing policy funds.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The dataset includes information on Toronto shelter system usage, such as daily occupancy rates, shelter types (e.g., for families, men, women), the number of residents, and available beds.

2. *How many instances are there in total (of each type, if appropriate)?*
	- Dataset is updated daily at 4 a.m., with new additions each day.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- No, the dataset is not a sample; it is the complete data directly uploaded from the shelters.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance now includes:
Date: A timestamp for when the data was recorded.
Occupancy Values: Two different columns for occupancy rates; some cells are missing values (NA).
Location Information: Specific details about the shelter's name and type.
Actual Beds/Occupancy: Counts for the actual number of beds available or occupied.
Sector: Categories such as "Youth," "Mixed Adult," "Men," or "Women" indicating the type of population served by the shelter.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- No explicit target or label is identified, but columns such as "Actual Beds" and "Occupancy" could be considered potential targets for analysis.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Yes, some cells contain "NA," indicating missing values in columns like occupancy rates or bed counts. The extent and pattern of missingness need to be evaluated further.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Relationships between instances are not explicitly marked. However, they can be inferred based on shared fields such as the location or sector.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- The dataset structure is not inherently split into different subsets (e.g., for training and testing). The data could be split based on shelter types, occupancy rates, or other factors for specific analyses.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- There is missing data in some columns, as shown by the "NA" values. This could be considered a source of noise or incompleteness. It's important to inspect the data further to check for other inconsistencies or redundancies.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained and does not link to or rely on any external resources.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, the dataset does not contain any confidential information.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No, the dataset does not contain any content that might be offensive or cause anxiety.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No, the dataset does not identify any specific sub-populations.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No, the dataset does not contain information that could be used to directly or indirectly identify individuals.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No, the dataset does not contain any sensitive information.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Data is typically reported daily by shelter staff and recorded in the city’s data systems to ensure the accuracy of the shelter census data.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Data is typically reported daily by shelter staff and recorded in the city’s data systems to ensure the accuracy of the shelter census data.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset is not a sample from a larger set; it is directly reported by shelter staff.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Shelter staff are responsible for data entry, compensation for shelter staff comes from their regular wages.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The dataset is updated daily.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- The dataset is provided and processed by official agencies of the City of Toronto.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data comes from the City of Toronto’s official open data website.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Data is reported daily by shelter staff and entered into the city's data systems.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- There is no involvement of individual personal data.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- The data may undergo cleaning and validation to ensure accuracy, which could include addressing missing data or correcting input errors.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	-  It is not specified whether the raw data is saved.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The availability of preprocessing software is not mentioned.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset is primarily used for policy-making, resource allocation, and monitoring the effectiveness of services for those experiencing homelessness. It may also be used to analyze trends and plan improvements to the shelter system.

2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No, there's no repository for papers or systems using the dataset.
3. *What (other) tasks could the dataset be used for?*
	- The dataset could be used for tasks such as analyzing shelter usage trends, predicting future shelter demand, assessing the impact of policy changes, and understanding the demographics of homelessness to improve targeted services.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No major concerns, but avoid misinterpreting macro trends for policy-making.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- Should not be used for individual-level analysis or assumptions about causes of homelessness.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The Shelter Census data is usually publicly available on the City of Toronto’s official website for use by the public, researchers, and policymakers.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Distributed via API.
3. *When will the dataset be distributed?*
	- Already publicly available.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The Housing and Homelessness Services department of the City of Toronto is responsible for maintaining the Shelter Census dataset.

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Email: 311@toronto.ca, Phone: 416-392-2489.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No additional information is provided.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- No additional information is provided.

\newpage

# References

### 2024-09-27 scripts/01-download_data.R
Update 01-download_data.R

@@ -23,5 +23,5 @@ marriage_licence_statistics <- marriage_licence_resources[5, ] %>%
   get_resource()
 analysis_data <- marriage_licence_statistics
 
-write.csv(analysis_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/raw_data.csv", row.names = FALSE)
+write.csv(analysis_data, "data/raw_data/raw_data.csv")
 

### 2024-09-27 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -12,7 +12,7 @@
 library(tidyverse)
 
 #### Cleaning the Dataset ####
-shelter_raw_data <- read_csv("C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/raw_data.csv")
+shelter_raw_data <- read_csv("data/raw_data/raw_data.csv")
 
 shelter_cleaned_data <-
   shelter_raw_data |>
@@ -26,4 +26,4 @@ shelter_cleaned_data <-
 
 
 #### Saving the Cleaned Dataset ####
-write.csv(shelter_cleaned_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data_clean.csv", row.names = FALSE)
+write.csv(shelter_cleaned_data, "data/analysis_data/analysis_data_clean.csv")

### 2024-09-27 scripts/03-test_data.R
Update 03-test_data.R

@@ -10,7 +10,7 @@
 #### Workspace setup ####
 library(tidyverse)
 
-data <- read_csv("C:\\Users\\EasonLi\\Downloads\\Toronto-main\\Toronto-main\\data\\analysis_data_clean.csv")
+data <- read_csv("data/raw_data/raw_data.csv")
 
 # test 1 date
 date_format_test <- all(grepl("^\\d{4}-\\d{2}-\\d{2}$", data$OCCUPANCY_DATE))

### 2024-09-27 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -24,4 +24,4 @@ simulated_data <- data.frame(
 head(simulated_data)
 
 # Save the simulated data to a CSV file
-write.csv(simulated_data, "C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/simulate_data/simulated_shelter_data.csv", row.names = FALSE)
+write.csv(simulated_data, "data/simulate_data/simulated_shelter_data.csv")

### 2024-09-27 paper/paper.qmd
Update and rename paper_Li.qmd to paper.qmd

--- 2024-09-27 paper/paper_Li.qmd
@@
 
-With respect to the structure of this paper, @sec-data includes an overview of the data and @sec-model show the results of model; @sec-discussion includes a discussion of the results; and @sec-appendix includes supplementary information.
+With respect to the structure of this paper, @sec-data includes an overview of the data and show the results of model; @sec-discussion includes a discussion of the results; and @sec-appendix includes supplementary information.
 
@@
 
-# Result {#sec-model}
+## analysis {#sec-data-analysis}
 
@@
   scale_color_manual(values = c("Bed Occupancy Rate" = "blue", "Room Occupancy Rate" = "red")) +
-  theme(legend.title = element_blank())
+  theme(legend.position = "top", legend.title = element_blank())
 ```

### 2024-09-27 paper/paper.qmd
Update and rename paper_Li.qmd to paper.qmd

@@
 
-The project is structure base on Rohan's principes [@rohan], and used the **R** programming language [@citeR], the `tidyverse` [@thereferencecanbewhatever] package were used to simulate the dataset and generate tests for it.
-The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@thereferencecanbewhatever] packages were then applied in order to download the raw Toronto Public shelter dataset.
-Next, the `tidyverse` package [@thereferencecanbewhatever] was used to clean the raw dataset and test the cleaned dataset.
+The project is structure base on Rohan's principes [@rohan], and used the **R** programming language [@citeR], the `tidyverse` [@citetidyverse] package were used to simulate the dataset and generate tests for it.
+The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@citetidyverse] packages were then applied in order to download the raw Toronto Public shelter dataset.
+Next, the `tidyverse` package [@citetidyverse] was used to clean the raw dataset and test the cleaned dataset.
+The `knitr` packages help illustrate the sample data (Xie 2024)
 
@@
 
-After loading the dataset using the R programming language [@citeR] and the `here` package [@citehere], the `tidyverse` [@thereferencecanbewhatever] package was used to generate graphs.
+After loading the dataset using the R programming language [@citeR] and the `here` package [@citehere], the `tidyverse` [@citetidyverse] package was used to generate graphs.
 In doing so, R code was adapted from [@tellingstorieswithdata].

### 2024-09-27 paper/paper_Li.qmd
Add files via upload

@@
 
-With respect to the structure of this paper, @sec-data includes an overview of the data and @sec-model show the results of model; @sec-discussion includes a discussion of the results; and @sec-appendix includes supplementary information.
+Concerning the structure of this paper, @sec-data includes an overview and visualization of the data; @sec-discussion includes a discussion of the results; and @sec-appendix includes supplementary information.
 
@@
 
-The dataset used in this analysis is from The "Daily Shelter & Overnight Service Usage" page , this website offers daily data of the occupancy and capacity of Toronto’s shelter system, including how many shelter beds are being used and the total available capacity, updated everyday except on weekends and holidays.
-The variables or measurements included in this analysis are "Bed Based Capacity", which refer to Occupancy is measured by the number of beds, typically for programs with shared sleeping areas; "Room Based Capacity" refer to ccupancy is measured by the number of rooms, usually for family programs or hotel-based programs where rooms are not shared between households.
+The dataset used in this analysis is from The "Daily Shelter & Overnight Service Usage" page, this website offers daily data on the occupancy and capacity of Toronto’s shelter system, including how many shelter beds are being used and the total available capacity, updated every day except on weekends and holidays.
+The variables or measurements included in this analysis are "Bed Based Capacity", which refers to Occupancy is measured by the number of beds, typically for programs with shared sleeping areas; "Room Based Capacity" to occupancy is measured by the number of rooms, usually for family programs or hotel-based programs where rooms are not shared between households.
 
-Another dataset called "toronto-shelter-system-flow" also includes data about Toronto shelters, but it is more focused on the characteristics of people living in shelters, not the macro data so I did not choose it.
+Another dataset called "toronto-shelter-system-flow" also includes data about Toronto shelters.
+Still, it is more focused on the characteristics of people living in shelters, not the macro data so I did not choose it.
 
-The project is structure base on Rohan's principes [@rohan], and used the **R** programming language [@citeR], the `tidyverse` [@thereferencecanbewhatever] package were used to simulate the dataset and generate tests for it.
-The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@thereferencecanbewhatever] packages were then applied in order to download the raw Toronto Public shelter dataset.
-Next, the `tidyverse` package [@thereferencecanbewhatever] was used to clean the raw dataset and test the cleaned dataset.
+The project is structure base on Rohan's principes [@rohan], and used the **R** programming language [@citeR], the `tidyverse` [@citetidyverse] package were used to simulate the dataset and generate tests for it.
+The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@citetidyverse] packages were then applied in order to download the raw Toronto Public shelter dataset.
+Next, the `tidyverse` package [@citetidyverse] was used to clean the raw dataset and test the cleaned dataset.
+The `knitr` packages help illustrate the sample data (Xie 2024)
 
@@
 
-@tbl-sample show a sample of cleaned occupancy data, Since the statistical record formats of different shelters are different, each row contains different information marked as NA, but this does not negattive affect the analysis.
+@tbl-sample show a sample of cleaned occupancy data, Since the statistical record formats of different shelters are different, each row contains different information marked as NA, but this does not negatively affect the analysis.
 \newpage
 
-# Result {#sec-model}
+## analysis {#sec-data-analysis}
 
@@
 Overall, the room occupancy rate (red) is higher and more stable, staying close to 100%.
-In contrast, the bed occupancy rate (blue) shows more fluctuations, generally ranging between 98% and 99%.
+In contrast, the bed occupancy rate (blue) shows more fluctuations, generally between 98% and 99%.
 
@@
 
-@fig-occupancy-capital-relat displays the average occupancy rate across various shelters, categorized by sector: Men, Women, Mixed Adult, and Youth.
+@fig-occupancy-capital-relat displays the average occupancy rate across various shelters, categorized by sector: Men, Women, Mixed Adults, and Youth.
 The X-axis represents the occupancy rate percentage (Occupancy Rate %), ranging from 0% to 100%, while the Y-axis lists the shelter locations (Location), showing the names of several shelters.
